---
title: "Methods and Initial Results"
author: "Julian McClellan"
date: "Due 5/17/17"
output:
  pdf_document:
    includes:
      in_header:
        header.tex
fontsize: 12pt
bibliography: citations.bib
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE)
library(tidyverse)
library(purrr)
library(lmtest)
library(lubridate)
library(stargazer)
library(ineq)
library(pROC)
library(modelr)
theme_set(theme_minimal())

dat_cmv_subs <- read_csv("changemyview/cmv_subs.csv") %>%
  mutate(OP_gave_delta = ifelse(OP_gave_delta == "True", TRUE, FALSE))
dat_cmv_auth_subs <- read_csv("changemyview/cmv_auth_subs.csv")

sumry_gave_delta <- dat_cmv_subs %>% 
  group_by(OP_gave_delta) %>%
  summarise(count = n()) %>%
  mutate(prop = count / sum(count))
```

## Model Justification

In the subreddit /r/changemyview, Reddit users (Redditors) post their views as
submissions on the subreddit, and other Redditors attempt to change these views.
If the Redditor who originally posted their view has changed their view to some
extent, they are able to award a "delta" to comment (and the Redditor who
authored that comment). These deltas are explicitly coded in the submission and
thus, one can say that for a given submission in /r/changemyview, whether or not
the original poster (OP) changed changed their mind at least once; i.e. they
awarded at least one delta.

Self-affirmation, a theory in psychology suggests:

> "[I]f global self-worth is temporarily bolstered by success in a second, unrelated domain, the individual should be more willing to tolerate a threat to the domain of interest."

[@correll2004affirmed]

In the context of submission in /r/changemyview, the "threats" to OP's domain of
interest are the comments that attempt to change their views. For many of the
OP's, their submissions to /r/changemyview are not their first submissions they
have made in Reddit. Any prior submissions an OP made can be considered
potential cases where his or her global self-worth may have been temporarily
bolstered. Looking at these submissions prior to an /r/changemyview submissions,
all of them receive a "score" of some number, the net of upvotes and downvotes
from other members of the community could serve as proxy for the "degree" to
which the individual's self-worth was bolstered. 

Additionally, one might argue that the very act of posting a submission bolsters
global self-worth on its own. Indeed, many psychological experiments testing
self-affirmation theory focus on global self-worth bolstered by the person
themselves [@cohen2000beliefs].

It might also be justified to also look at the *total* score of the Redditor's prior submissions, to explicitly see the lifetime score of their submissions prior to that /r/changemyview submission. In a generalized linear model, we can account for this with an interaction term between the number of submissions and the average score per submission.

Thankfully, the entire submission history of any Redditor is publically available, so all of these statistics can be gathered and calculated. 

### Past Submission Diversity

Additionally, one might also want to test for an sort of "echo" chamber effect.
The subreddit in which submissions were made can be obtained, and there might 
be a difference in how likely a Redditor is to award a "delta" if that
Redditor's past submissions took place across a relatively diverse set of
subreddits versus a select few. Gini index can serve as suitable metric for this 
measure.

## Model Definition

The binary nature of our outcome variable, whether or not OP awards a delta, makes the following logistic model viable:

$$prob\_OP\_award\_delta = \frac{1}{1 + exp(-(\beta_0 + \beta_{1}X_1+
\beta_{2}X_2 + \beta_{3}X_3 + \beta_{4}X_1X_2))}$$

where: 
$$\left\{\begin{matrix}
X_1 & = & num\_prior\_subs& \\ 
X_2 &  =&avg\_sub\_score \\ 
X_3 &  =& subreddit\_gini\_index
\end{matrix}\right.$$

```{r make_model_dat, eval=FALSE}
model_indep_vars <- function(submission, dat_cmv_auth_subs){
  force(dat_cmv_auth_subs)
  prior_subs <- dat_cmv_auth_subs %>%
    filter(author == submission$author,
           created_utc < submission$created_utc)
  
  if (nrow(prior_subs) == 0){
    has_priors <- FALSE
  } else{
    has_priors <- TRUE
  }
  
  subreddit_dist <- prior_subs %>%
    group_by(subreddit) %>%
    summarise(n = n())
  
  # browser()
  
  (sub_info <- c(nrow(prior_subs),
                mean(prior_subs$score),
                ineq(subreddit_dist$n, type = "Gini"),
                has_priors))
}

model_dat <- dat_cmv_subs %>%
  by_row(model_indep_vars, dat_cmv_auth_subs = dat_cmv_auth_subs, 
         .collate = "cols", .to = "indep_var") %>%
  rename(num_prior_subs = indep_var1,
         avg_sub_score = indep_var2,
         gini_index = indep_var3,
         has_priors = indep_var4) %>%
  filter(has_priors == TRUE) %>%
  select(OP_gave_delta, num_prior_subs, avg_sub_score, gini_index)

write_csv(model_dat, "model_dat.csv")
```

```{r read_model_dat}
model_dat <- read_csv("model_dat.csv")
indep_vars <- model_dat %>%
  select(-OP_gave_delta)

sumry_model_gave_delta <- model_dat %>% 
  group_by(OP_gave_delta) %>%
  summarise(count = n()) %>%
  mutate(prop = count / sum(count))
```

## Data 

The data includes all `r nrow(dat_cmv_subs)` submissions made by `r length( unique(dat_cmv_subs$author))` unique authors. Of these `r nrow(dat_cmv_subs)`
submissions, `r sumry_gave_delta$count[1]` of them, or 
`r paste(round(sumry_gave_delta$prop[1] * 100, 2), "%")` did not have their OP
award a delta. (The view of the OP was not changed). The other `r sumry_gave_delta$count[2]` of the posts, or `r paste(round(sumry_gave_delta$prop[1] * 100, 2), "%")` had their OP award a delta.

From this initial set of submissions, we are only interested in submissions in
which the OP had prior submission history.

From the initial `r nrow(dat_cmv_subs)` submissions, only `r nrow(model_dat)` (`r paste(round(100 * nrow(model_dat) / nrow(dat_cmv_subs), 2), "%")`) were made by
OP's with prior submission history. Of these `r nrow(model_dat)`
submissions, `r sumry_model_gave_delta$count[1]` of them (`r paste(round(sumry_model_gave_delta$prop[1] * 100, 2), "%")`) did not have
their OP award a delta. (The view of the OP was not changed). The other `r sumry_model_gave_delta$count[2]` of the posts (`r paste(round(sumry_model_gave_delta$prop[1] * 100, 2), "%")`) of had their OP
award a delta.

A summary table for:  

```{r, results="asis"}
stargazer(as.data.frame(indep_vars), 
          title = "Independent Variable Summary Statistics",
          header = FALSE, type = "latex", 
          summary.stat =  c("mean", "sd", "min", "p25", "median", "p75", "max")) 
```

## Initial Results

Now we compute the logistic model as described above:

$$prob\_OP\_award\_delta = \frac{1}{1 + exp(-(\beta_0 + \beta_{1}X_1+
\beta_{2}X_2 + \beta_{3}X_3 + \beta_{4}X_1X_2))}$$

where: 
$$\left\{\begin{matrix}
X_1 & = & num\_prior\_subs& \\ 
X_2 &  =&avg\_sub\_score \\ 
X_3 &  =& subreddit\_gini\_index
\end{matrix}\right.$$

---

The table below shows statistical signifiance at the `0.01` level for the number
of prior submissions and the total prior submissions score we created as an
interaction term. Now let's evaluate the model accuracy with a look at the area
under the ROC curve (AUC).

```{r logistic_results, results="asis"}
logit_init <- glm(OP_gave_delta ~ num_prior_subs + avg_sub_score + num_prior_subs * avg_sub_score + gini_index,
                  data = model_dat)
stargazer(logit_init, title = "Initial Logistic Regression Results", 
          covariate.labels = c("Number of prior Submissions", 
                            "Average prior submission score",
                            "Gini Index",
                            "Total Prior Submission Score (Interaction)"),
          keep.stat = c("rsq", "ser", "wald", "adj.rsq", "rsq"),
          type = "latex",
          header = FALSE)

logit2prob <- function(x){
  exp(x) / (1 + exp(x))
}
```


```{r logistic_auc}
delta_accuracy <- model_dat %>%
  add_predictions(logit_init) %>%
  mutate(pred = logit2prob(pred),
         prob = pred,
         pred = as.numeric(pred > .5))

auc_delta <- auc(delta_accuracy$OP_gave_delta, delta_accuracy$prob)

plot(roc(delta_accuracy$OP_gave_delta, delta_accuracy$prob), 
     main =
       sprintf("ROC Curve for Initial Logistic Model | AUC = %.3f", auc_delta))
```

Although we only get an AUC of `r auc_delta` this is still very promising. Tan
et al. [-@tan2016winning] were only able to get an AUC of `.54`, by looking at a
large number of features extracted from the langauge of the CMV submission
itself. In the logistic model given above, we use much more primitive features,
simply looking at the OP's past submissions, their scores, and a simple measure
of subreddit diversity in these submissions. Additionally, Tan et al. utilized a filtering
procedure, based on OP interaction and the number of unique respondents, to remove
submissions that did not have enough debate (in their view) to warrant a delta; no such
filtering procedure has been used here.

The statistical signifiance of some of our features provides some evidence that
self-affirmation plays a role in whether or not the original poster of
/r/changemyview submission will change their view (award a delta). Additionally,
because the original text of each individual submission is also available for
natural language processing, this logistic model can be improved by adding these
sorts of language features as well. For instance, Tan et al. utilized first
person pronoun useage as a proxy for self-affirmation in the /r/changemyview submission, and first-person pronoun useage in prior submission content can be be extracted as well.

\newpage 

## References