---
title: "Perspectives Research Problem Set 1"
author: "Julian McClellan"
date: "Due 4/19/17"
output: github_document
bibliography: citations.bib
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE)
library(tidyverse)
library(ggplot2)
library(stargazer)
library(lubridate)

theme_set(theme_minimal())
food_inspections <- read_csv('~/Downloads/Food_Inspections.csv') %>%
  dplyr::select(-c(Location, Address, `Inspection ID`, State)) %>%
  mutate(`Inspection Date` = mdy(`Inspection Date`),
         Zip = factor(Zip),
         Risk = gsub("[^1-3]", "", Risk),
         Risk = as.numeric(Risk),
         City = toupper(City))
num_obs <- nrow(food_inspections)
```

# Data

The data is drawn from inspections of restaurants and other food establishments in Chicago from January 1, 2010 to the present
(April 13, 2017).

Inspections are performed by staff from the Chicago Department of Public Health’s Food Protection Program using a standardized
procedure. The results of the inspection are inputted into a database, then reviewed and approved by a State of Illinois Licensed
Environmental Health Practitioner. The data is available on [Chicago's open data portal](https://data.cityofchicago.org/Health-Human-Services/Food-Inspections/4ijn-s7e5/data).

At a meeting of the American Public Health Association, [@schenk2015food] described an [open source project](https://github.com/Chicago/food-inspections-evaluation) in which:

> [T]he City of Chicago’s Department of Innovation and Technology (DoIT), in collaboration with an insurance company, and the CDPH, together developed advanced analytics to forecast 
> food establishments that are most likely to have critical violations, which are most likely to contribute to food borne 
> illness, so that they may be inspected first.

Making a *single* descriptive statistics table isn't exactly the most helpful thing
given that the majority of the variables of categorical with differing numbers of categories, but here you go:

```{r tables, results='asis'}
licence_mipu <- food_inspections %>%
  group_by(`License #`) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()
  
lname_mipu <- food_inspections %>%
  group_by(`DBA Name`) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()

pname_mipu <- food_inspections %>%
  group_by(`AKA Name`) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()

ftype_mipu <- food_inspections %>%
  group_by(`Facility Type`) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()

risk_mipu <- food_inspections %>%
  group_by(Risk) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()

city_mipu <- food_inspections %>%
  group_by(City) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()

zip_mipu <- food_inspections %>%
  group_by(Zip) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()

itype_mipu <- food_inspections %>%
  group_by(`Inspection Type`) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()

results_mipu <- food_inspections %>%
  group_by(Results) %>%
  summarise(n = n()) %>%
  {.$n} %>%
  median()

calc_mode <- function(vector, get_prop = FALSE){
  if (get_prop){
  sorted <- sort(table(vector), decreasing = TRUE)  
  prop <- sorted[1] / sum(sorted)
  sprintf("%.3f%%", 100 * prop)
  } else {
  names(sort(table(vector), decreasing = TRUE))[1] 
  }
}

 calc_anti_mode <- function(vector, get_prop = FALSE){
  if (get_prop){
  sorted <- sort(table(vector), decreasing = FALSE)  
  prop <- sorted[1] / sum(sorted)
  sprintf("%.3f%%", 100 * prop)
  } else {
  names(sort(table(vector), decreasing = FALSE))[1] 
  }
 }
 
 
attach(food_inspections)
cat_table <- tibble(`Categorical Variables` =
                           c("License #", "Legal Name", "Public Name",
                             "Facility Type", "Risk (lower = riskier)",
                             "City", "Zip Code", "Inspection Type",
                             "Inspection Results"),
                   `unique Values` = c(length(unique(`License #`)),
                                                length(unique(`DBA Name`)),
                                                length(unique(`AKA Name`)),
                                                length(unique(`Facility Type`)),
                                                length(unique(Risk)),
                                         length(unique(City)), length(unique(Zip)), 
                                                length(unique(`Inspection Type`)),
                                                length(unique(Results))),
                         `Median inspections per unique value` = c(
                           licence_mipu, lname_mipu, pname_mipu, ftype_mipu,
                           risk_mipu, city_mipu, zip_mipu, itype_mipu, results_mipu),
                   `Most Frequent Value` = c(calc_mode(`License #`),
                                             calc_mode(`DBA Name`),
                                             calc_mode(`AKA Name`),
                                             calc_mode(`Facility Type`),
                                             calc_mode(Risk),
                                             calc_mode(City),
                                             calc_mode(Zip),
                                             calc_mode(`Inspection Type`),
                                             calc_mode(Results)
                                             ),
                   `Most Frequent Value Proportion` = c(calc_mode(`License #`, TRUE),
                                             calc_mode(`DBA Name`, TRUE),
                                             calc_mode(`AKA Name`, TRUE),
                                             calc_mode(`Facility Type`, TRUE),
                                             calc_mode(Risk, TRUE),
                                             calc_mode(City, TRUE),
                                             calc_mode(Zip, TRUE),
                                             calc_mode(`Inspection Type`, TRUE),
                                             calc_mode(Results, TRUE)
                                             ),
                   `Least Frequent Value` = c(calc_anti_mode(`License #`),
                                             calc_anti_mode(`DBA Name`),
                                             calc_anti_mode(`AKA Name`),
                                             calc_anti_mode(`Facility Type`),
                                             calc_anti_mode(Risk),
                                             calc_anti_mode(City),
                                             calc_anti_mode(Zip),
                                             calc_anti_mode(`Inspection Type`),
                                             calc_anti_mode(Results)
                                             ),
                   `Least Frequent Value Proportion` = c(calc_anti_mode(`License #`, TRUE),
                                             calc_anti_mode(`DBA Name`, TRUE),
                                             calc_anti_mode(`AKA Name`, TRUE),
                                             calc_anti_mode(`Facility Type`, TRUE),
                                             calc_anti_mode(Risk, TRUE),
                                             calc_anti_mode(City, TRUE),
                                             calc_anti_mode(Zip, TRUE),
                                             calc_anti_mode(`Inspection Type`, TRUE),
                                             calc_anti_mode(Results, TRUE)
                                             )
                   )
```

```{r visualization, results='asis'}
```

```{r slice, results='asis'}