---
title: "Perspectives Research Problem Set 1"
author: "Julian McClellan"
date: "Due 4/19/17"
output: github_document
bibliography: citations.bib
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE)
library(tidyverse)
library(ggplot2)
library(stargazer)
theme_set(theme_minimal())
food_inspections <- read_csv('~/Downloads/Food_Inspections.csv')
num_obs <- nrow(food_inspections) %>%
  mutate(State = "IL")

food_inspections <- food_inspections %>% 
  na.omit()
num_full <- nrow(food_inspections)
num_obs_na <- num_obs


unique_dba <- length(unique(food_inspections$`DBA Name`))
unique_aka <- length
  
stargazer(as.data.frame(food_inspections), type = 'html', median = TRUE, mean.sd = FALSE)
```

# Data

The data is drawn from inspections of restaurants and other food establishments in Chicago from January 1, 2010 to the present
(April 13, 2017).

Inspections are performed by staff from the Chicago Department of Public Health’s Food Protection Program using a standardized
procedure. The results of the inspection are inputted into a database, then reviewed and approved by a State of Illinois Licensed
Environmental Health Practitioner. The data is available on [Chicago's open data portal](https://data.cityofchicago.org/Health-Human-Services/Food-Inspections/4ijn-s7e5/data).

At a meeting of the American Public Health Association, [@schenk2015food] described an [open source project](https://github.com/Chicago/food-inspections-evaluation) in which:

> [T]he City of Chicago’s Department of Innovation and Technology (DoIT), in collaboration with an insurance company, and the CDPH, together developed advanced analytics to forecast 
> food establishments that are most likely to have critical violations, which are most likely to contribute to food borne 
> illness, so that they may be inspected first.

Summary statistics of the data follow below:

```{r, results='asis'}
```